### 2017-08-16 doc/MSc-dissertation.tex
results

\usepackage{graphics}
\usepackage{listings}
\usepackage{fullpage,epsf,graphicx, amstext,url,bm,amsmath, amssymb}

@@

\section{Dynamo action}

@@

\section{DNS resolution}

@@

Another important part of the eDNS2017 code is the type of forcing used to inject energy into the system. It is worth noting that the code allows to study isotropic turbulence for decaying (no energy injected into the system) or forcing (energy injected in order to reach a steady state)~\cite{LinkmannMoritzFrederikLeon2016Spim}. This forcing comes as $\bm f_u$ and $\bm f_b$ in eqns.~\ref{eq3.1} and~\ref{eq3.2}. However, for the simulations that were performed only the velocity field was forced, and from it energy was transferred to the magnetic field through the Lorentz force. From stationary simulations, the forcing function used injects energy by feeding a rescaling of the velocity field into the large spatial scales. More precisely, $\bm f_u$ is given by:

@@

A detailed account of the simulations that were generated, along with their motivations, is provided here. The first simulation was done in a large box, $N^3=1024^3$, for which viscosity ($\nu=0.005$) was ten times larger than magnetic diffusivity, i.e. $Pr_M=\nu/\eta=10$. As it was mentioned before, the magnetic Prandtl number compares the kinetic and the magnetic channel. That is, it compares how turbulent the magnetic field is relative to the velocity field, which is equivalent of comparing how much energy is dissipated through the magnetic channel against the kinetic channel.

For the simulation with $N^3=1024^3$ grid points, \texttt{Paraview}, a slice of the velocity and the magnetic fields' magnitude was plotted. Figure~\ref{fig5.1} shows this slice, where the magnetic field is clearly more fine-grained than the velocity field. Which is not surprising since the Kolmogorov microscales for this system are $k_\nu = 33$ and $k_\eta = 185$.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\textwidth]{{img/f_1024_0.005_10_fieldslice}.png}
\caption{Heat map of $\vert u \vert$ (left) and $\vert b \vert$ (right) of a vertical  slice of a $N^3=1024^3$ box ~\cite{schekochihin2004simulations}.}
\label{fig5.1}
\end{figure}

@@

in~\cite{haugen2004simulations} they mention that the Kazantsev $k^{3/2}$ spectrum should appear when the magnetic field is strong enough and in the kinetic regime (small enough wave numbers such that nonlinear term is unimportant). This happens in the highPrM runs! a plot of it is in {highPrM/img/Kazantsev} for $\eta=0.01$. It is comparable with figure 4(a) in~\cite{schekochihin2004simulations}

@@

This section should be written in standard scientific
language. Standard techniques in your research field should not be
written out in detail. In computational projects this section should
be used to explain the algorithms used and the layout of the
computational code. A copy of the actual code may be given in the
appendices if appropriate.

This section should emphasise the philosophy of the approach used and
detail novel techniques. However please note: this section should not
be a blow-by-blow account of what you did throughout the project. It
should not contain large detailed sections about things you tried and
found to be completely wrong! However, if you find that a technique
that was expected to work failed, that is a valid result and should be
included.

Here logical structure is particularly important, and you may find
that to maintain good structure you may have to present the
explorations/calculations/computations/whatever in a different order
from the one in which you carried them out.

>>>>

\usepackage{graphics}
\usepackage{listings}
\usepackage{fullpage,epsf,graphicx, amstext,url,bm,amsmath, amssymb, caption}

@@

\section{Dynamo action}
\label{sec3.3}

@@

\section{DNS resolution}
\label{sec4.2}

@@

Another important part of the eDNS2017 code is the type of forcing used to inject energy into the system. It is worth noting that the code allows to study isotropic turbulence for decaying (no energy injected into the system) or forcing (energy injected in order to reach a steady state)~\cite{LinkmannMoritzFrederikLeon2016Spim}. This forcing comes as $\bm f_u$ and $\bm f_b$ in eqns.~\ref{eq3.1} and~\ref{eq3.2}. However, for the simulations that were performed only the velocity field was forced, and from it energy was transferred to the magnetic field through the Lorentz force. From stationary simulations, the forcing function used\footnote{The forcing used for the non-helical simulations was negative damping, which is thoroughly described in~\cite{chalupa2017comparison}} injects energy by feeding a rescaling of the velocity field into the large spatial scales. More precisely, $\bm f_u$ is given by:

@@

A detailed account of the simulations that were generated, along with their motivations, is provided in this chapter. The first section obtains the results from a large forced non-helical simulation, with $N^3=1024^3$ grid points, and the resulting characteristic features it has. A set of remarks regarding the magnetic Prandtl number are given as well. Afterwards, a series of forced non-helical simulations are used to show that the kinetic energy spectrum follows the Kolmogorov $k^{-5/3}$ scaling when the magnetic channel is weak (in comparison with the kinetic channel). The same is done to check for Kazantsev $k^{3/2}$ scaling for the magnetic energy spectrum in simulations with $Pr_M \geq 1$ that have achieved dynamo action~\cite{haugen2004simulations}.

After these brief consistency checks for the eDNS2017 code, several simulations were used in order to study the onset of dynamo action (all were forced with the negative damping forcing described in~\cite{chalupa2017comparison} and, hence, without injection of helicity). The definition of dynamo action that was worked with was that of~\cite{schnack2009lectures} (page 266), which was restated here in section~\ref{sec3.3} and arguing that when simulations reach a steady state in forced simulations it can be considered as if an ``infinite'' time was achieved. Which is supported by the ergodic theorem in~\cite{biskamp1997nonlinear}. Furthermore, it is claimed that from these simulations, all those with dynamo action are undergoing small-scale dynamo action since they have no kinetic helicity. At first, and following from claims of~\cite{schekochihin2004critical} about the critical magnetic Reynolds number having a dependence of the Reynolds number, a critical magnetic Reynolds number was found for $\eta=0.01 \text{ and } 0.0125$ that showed dynamo action at $Pr_M=1$ by decreasing $Pr_M$, and then by increasing it to $Pr_M > 1$. Doing so it was found that there were two different $Re_M^{crit}$ for each value of $\eta$ depending on whether $Pr_M \leq 1$ or $Pr_M \geq 1$.

Finally, a few remarks regarding the injection of kinetic helicity (through helical forcing) into different systems is provided. Namely, once helicity is injected into the system the critical magnetic Reynolds number (both with $Pr_M < 1$ and $Pr_M > 1$) decreases. This boost of the total magnetic energy is likely due to $\alpha$-effect, which is characteristic to large-scale dynamo action. For this reason it is possible to say that the injection of kinetic helicity leads to the generation of large-scale dynamos, which are known to generate mean magnetic fields.

\section{Simulation with $N^3=1024^3$ grid points}

The first simulation was done in a large box, $N^3=1024^3$, for which viscosity ($\nu=0.005$) was ten times larger than magnetic diffusivity, i.e. $Pr_M=\nu/\eta=10$. As it was mentioned before, the magnetic Prandtl number compares the kinetic and the magnetic channels. That is, it compares how turbulent the magnetic field is relative to the velocity field, which is equivalent to comparing the magnetic Reynolds number against the kinetic magnetic Reynolds number $Pr_M=Re_M/Re$.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\textwidth]{{img/f_1024_0.005_10_fieldslice}.png}
\caption{Heat map of $\vert u \vert$ (left) and $\vert b \vert$ (right) of a vertical  slice of a $N^3=1024^3$ box.}
\label{fig5.1}
\end{figure}

For this simulation, \texttt{Paraview} was used to create a two-dimensional slice of the velocity and the magnetic fields' magnitude. Figure~\ref{fig5.1} shows these slices, and a few comments can be made about them. First of all, it is noted that the velocity field has a much stronger magnitude than the magnetic field. This is likely due to the fact that the velocity field is the one that is forced with the negative damping forcing at $k_f=2$. It is also interesting to note that there appears to be a sort of duality between $\vert u \vert$ and $\vert b \vert$ - which is apparent by the structures formed by each field. For example, the bright red area in the top central section of the $\vert b \vert $ image has an inverse dual on the $\vert u \vert$ slice. This inverse duality can be seen all over fig.~\ref{fig5.1}.

Additionally, the magnetic field is clearly more fine-grained than the velocity field, which makes sense given that the $\bm b$ is ten times more turbulent than $\bm u$. Quantitatively, this can be interpreted through the kinetic and magnetic Reynolds numbers. As it was mentioned before, these are a measure of how turbulent a specific field is. For the magnetic field, $Re_M=1291\pm50$ and for the velocity field, $Re=129\pm5$, where the errors obtained come from the errors output by the \texttt{time\_av.sh} post-processing script along with basic (addition, division, powers) error propagation rules (eqns.~\ref{eq5.1}). This difference in Reynolds numbers comes with an important difference of the dissipation wave numbers of each field.

\begin{subequations}
 \begin{align}
  Q &= a + b - x - y \nonumber \\
  \delta Q &= \sqrt{(\delta a)^2 + (\delta b)^2 + (\delta x)^2 + (\delta y )^2} \label{eq5.1.1} \\
  Q &= \frac{a b}{x y} \nonumber \\
  \delta Q &= \vert Q \vert \sqrt{\left( \frac{\delta a}{a} \right)^2 + \left( \frac{\delta b}{b} \right)^2 + \left( \frac{\delta x}{x} \right)^2 + \left( \frac{\delta y}{y} \right)^2} \label{eq5.1.2} \\
  Q &= x^n \nonumber \\
  \delta Q &= \vert Q \vert \vert n \vert \frac{\delta x}{\vert x \vert } \label{eq5.1.3}
 \end{align}
\label{eq5.1}
\end{subequations}

Given that laminar flows have a smaller dissipation wave numbers ($k_\nu$ and $k_\eta$) than turbulent flows, and that energy dissipation through either channel mainly occurs close to the respective dissipation wave numbers, the magnetic energy spectrum should have more energy in the larger wave numbers than the kinetic energy spectrum (since $k_\nu < k_\eta$). In fact, these dissipation wave numbers can be obtained through eqns.~\ref{eq2.7} and~\ref{eq3.5}, and correspond to $k_\nu = 33$ and $k_\eta=183$. Using a simple \texttt{bash} script, the spectral image of the $N^3=1024^3$ box's kinetic and magnetic energies is created, and shown in fig.~\ref{fig5.2}, and confirms the aforementioned suspicion.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\textwidth]{{img/1024/spectra_1024_0.005_10}.png}
\caption{Kinetic (solid red line) and magnetic (dashed blue line) energy spectra. The dissipation wave numbers $k_\nu=33$ and $k_\eta=183$ are also shown. $k=340$ is the maximum wave number that the $N^3=1024^3$ box allows.}
\label{fig5.2}
\end{figure}

In fig.~\ref{fig5.2} the kinetic energy spectrum ($E_K(k)$) is only more energetic in the smaller wave numbers (largest eddies), and decays quickly for the large ones (smaller length scales). In particular, the maximum value of $E_K(k)$ is at the forcing wave number $k_f=2$, and decays as $k$ increases. On the other hand, $E_M(k)$ is maximally energetic around $k\sim10$, and remains at a high energy for a broad range of wave numbers until $k$ is close to the magnetic diffusive wave number $k_\eta$.

Another important qualitative description that can be obtained regards the energy dissipation through each field. Namely, since the magnetic Prandtl number compares the Reynolds numbers of both fields $Pr_M=Re_M/Re$, which is a comparison the energy dissipation of each channel (since the $k_\nu$ and $k_\eta$ are intimately related to energy dissipation)~\cite{brandenburg2011dissipation}. As a result, for the simulation of $N^3=1024^3$ grid points with $Pr_M=10$, one would expect that the energy dissipation through the magnetic channel to be larger than that of the kinetic channel, which is exactly what fig.~\ref{fig5.3} shows.

\begin{figure}[!ht]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \captionsetup{justification=centering}
  \includegraphics[width=0.95\linewidth]{{img/1024/energy_dissipation_1024_0.005_10}.png}
  \caption{Time evolution of the total magnetic (dashed blue line) and kinetic (solid red line) energy dissipations for the simulation with $N^3=1024^3$ grid points.}
  \label{fig5.3}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \captionsetup{justification=centering}
  \includegraphics[width=0.95\linewidth]{{img/1024/energy_1024_0.005_10}.png}
  \caption{Time evolution of the total magnetic (dashed blue line) and kinetic (solid red line) energies for the simulation with $N^3=1024^3$ grid points.}
  \label{fig5.4}
\end{minipage}
\end{figure}

Finally, and looking at fig.~\ref{fig5.4} that plots the total magnetic (dashed blue line) and kinetic (solid red line) energies in time, it is possible to say that, if a steady state has been reached, dynamo action has been sustained. After stating that the code-time interval [40:50] is steady state in nature, ensemble averages can be obtained. This is due to ergodicity, taking time averages can be associated with taking ensemble averages~\cite{biskamp1997nonlinear}. As a consequence, it is possible to say that, since the average total magnetic energy at the end of various simulations is not zero, then dynamo action has been reached. Since there is no kinetic helicity in the system the dynamo action associated to this system can be said to be small-scale dynamo action.

\subsection{Simulation resolution}

Section~\ref{sec4.2} strongly emphasised the importance of the resolution required for a given simulation. A few additional remarks about it are presented here. Strictly on physical grounds it is straight forward to see that in order to properly resolve a simulation it is paramount to reach the wave numbers that dissipate most energy in the system. Making a simulation with the same parameters as the $N^3=1024^3$ simulation ($\nu=0.005$, $Pr_M=10$), but with a considerably smaller box, $N^3=128^3$ it is possible to see how this affects the results obtained.

It is worth mentioning that the viscous dissipation wave number is $k\nu=33$ and that the maximum wave number that a $N^3=128^3$ box can achieve is $k_{max}=41$. As a result, one would expect the kinetic channels to look pretty similar in both simulations. However, since the magnetic dissipation wave number is $k_\eta=183$, which is considerably larger than the maximum wave number reached in the $N^3=128^3$ box, it is clear that there should be a substantial difference in the time evolution of the magnetic energy dissipation of both box sizes. Figure~\ref{fig5.5} shows the difference in the kinetic and magnetic energy dissipations of both boxes and fig.~\ref{fig5.6} similarly for $E_K(t)$ and $E_M(t)$.

\begin{figure}[!ht]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \captionsetup{justification=centering}
  \includegraphics[width=0.95\linewidth]{{img/resolution/energy_dissipation_1024_0.005_10}.png}
  \caption{Time evolution of the total magnetic (dashed blue line $N^3=1024^3$ and dashed dark blue line  $N^3=128^3$) and kinetic (solid red line $N^3=1024^3$ and dashed dark red line  $N^3=128^3$) energy dissipations.}
  \label{fig5.5}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \captionsetup{justification=centering}
  \includegraphics[width=0.95\linewidth]{{img/resolution/energy_1024_0.005_10}.png}
  \caption{Time evolution of the total magnetic (dashed blue line $N^3=1024^3$ and dashed dark blue line  $N^3=128^3$) and kinetic (solid red line $N^3=1024^3$ and dashed dark red line  $N^3=128^3$) energies.}
  \label{fig5.6}
\end{minipage}
\end{figure}

Given that the dissipation of energy mainly occurs close to the dissipative wave numbers, it is no surprise to see that the simulation that was not large enough to resolve $k_\eta=183$ shows considerably less energy dissipation in the magnetic channel (dashed dark blue and blue lines). It is also straight forward to understand why, on the other hand, the kinetic energy dissipation looks roughly the same - since the viscous energy dissipation was resolved in both simulations, most of the kinetic energy dissipation was appropriately calculated.

Figure~\ref{fig5.6} clearly shows that the magnetic energy that is calculated by the simulation with smaller box size is weaker than that obtained with $N^3=1024^3$ grid points. This poses an important problem when trying to determine which simulations maintain dynamo action. The reason behind this is that dynamo action is said to be maintained when the magnetic energy at the end of the simulation is nonzero. However, if the simulation that is being considered lacks the required spatial resolution such that the magnetic energy is (mistakenly) calculated to be weaker (particularly, $E_M(t=t_f)=0$) than its real value ($E_M(t=t_f)=0$), then it could be incorrectly concluded that no dynamo action is sustained by the given values.

If a similar problem happens when calculating the growth rate in decaying simulations, then the results of~\cite{schekochihin2004critical} can be criticised. The main reason to claim this is because all their remarks are given for simulations with $\eta=0.001$ (and some with $\nu$ as small as $0.000125$), which in the eDNS2017 code would have to be resolved by (at least)\footnote{For a simulation generated with eDNS2017 with $\nu=0.001$, the associated dissipative wave number is $k_\nu=103$. $N^3=128^3$ simulations have a maximum wave number $k_{max}^{128^3}=41$ and for a $N^3=256^3$ it is $k_{max}^{256^3}=84$. Consequently, a box size of at least $N^3=512^3$ would be required to resolve a simulation with $\nu=0.001$ (assuming $\eta \geq \nu$), since $k_{max}^{512^3}=169$.} a box size of of $N^3=512^3$ and, for $\nu=0.000125$, $N^3=1024^3$ grid points (or larger)\footnote{For the $N^3=1024^3$ simulation generated ($\nu=0.005$ and $\eta=0.0005$), $k_\eta=183$. $N^3=1024^3$ simulations have a maximum wave number $k_{max}^{256^3}=340$. Consequently, a box size of at least $N^3=1024^3$ would be required to resolve a simulation with $\eta=0.0005$ (assuming $\nu \geq \eta$). Thus it is possible that for a simulation with $\nu=0.000125$ a box even larger than $N^3=1024^3$ would need to be used since $k_{\nu, \eta}$ increase as $\nu, \eta$ decrease.}. This would mean that the simulation performed in~\cite{schekochihin2004critical} are under resolved. And, as a result, some simulations which they claim to have not achieved dynamo action might actually have.  However, a different behaviour could be seen for the growth rate in decaying simulations, in which case the simulations in~\cite{schekochihin2004critical} could have been adequately generated (than that for the magnetic energy). Nonetheless, if the box sizes used in~\cite{schekochihin2004critical} would have been replicated in the forced non-helical simulations generated with the eDNS2017 code for the aforementioned parameters, then all simulations would not have encompassed the entire physical picture.

Finally, the kinetic and magnetic energy spectra were also plotted and, surprisingly, found to be very similar.

\section{Kolmogorov and Kazantsev scales}

A well known theoretical prediction regarding the spectral energy cascade in hydrodynamics is the so-called Kolmogorov scaling (section~\ref{sec2.4}). Namely, Kolmogorov's theory focuses on the inertial subrange, which is close to the Taylor length scale ($\lambda_u=\sqrt{15\nu/\varepsilon}u_{rms}=\sqrt{10\nu E_K/\varepsilon}$) and large enough that dissipative effects due to viscosity do not affect the fluid motion and small enough that the fluid's large scale motions are also neglected.

@@

in~\cite{haugen2004simulations} they mention that the Kazantsev $k^{3/2}$ spectrum should appear when the magnetic field is strong enough and in the kinetic regime (small enough wave numbers such that nonlinear term is unimportant). This happens in the highPrM runs! a plot of it is in {highPrM/img/Kazantsev} for $\eta=0.01$. It is comparable with figure 4(a) in~\cite{schekochihin2004simulations} x

@@
